Introducing the \NAME Framework\\ for Efficient Intrusion Detection\\ in the Internet of Things

### 2014-10-14 paper.tex
added subsection on the DSL component

% additional packages and utility commands
\usepackage{minted}
\usepackage{flushend}

@@

@@

\begin{figure}[H]
\begin{algorithmic}[1]
  \Let{$msg$}{$network$} \Comment{all observed messages}
  \ForEach{$algorithm \in algorithms$}
    \State \Call{algorithm.progress\_message}{msg}
  \EndFor
  \State \dots
  \Comment{at a given interval}
  \ForEach{$algorithm \in algorithms$}
    \State \Call{algorithm.do\_housekeeping}{}
  \EndFor
\end{algorithmic}
\captionof{listing}{Sequential implementation of several ID algorithms\label{alg:id-algo-application}}
\end{figure}

@@

\begin{figure}[H]
\begin{algorithmic}[1]
  \Require{nodes, global storage for information about nodes}
  \Function{process\_message}{$msg$}
    \ForEach{$byte \in msg$} \label{alg:id-algo-pattern-loop1}
     \State \dots \Comment{analyse byte-sequences}
    \EndFor
    \Let{$nodes_x$}{value}  \Comment{optionally update node info}
    \State \Call{send}{$nodes_y$, ``info''} \Comment{optionally exchange info} \label{alg:id-algo-pattern-send1}
  \EndFunction
  \State
  \Function{do\_housekeeping}{}
    \ForEach{$node \in nodes$} \label{alg:id-algo-pattern-loop2} \label{alg:id-algo-pattern-common-data}
      \If{$node > \dots$} \Comment{validate recorded value}
        \State \dots \Comment{take actions}
        \State \Call{send}{$node$, ``info''} \Comment{e.g. communicate} \label{alg:id-algo-pattern-send2}
      \EndIf
    \EndFor
  \EndFunction
  \State
\end{algorithmic}
\captionof{listing}{ID Algorithm Pattern\label{alg:id-algo-pattern}}
\end{figure}

@@

>>>>

% additional packages and utility commands
\newcommand{\TODO}{\textbf{\color{red}TODO}}

\usepackage{minted}
\usepackage{flushend}

@@

\usepackage{listings}

@@

\begin{figure}[H]
\captionof{lstlisting}{Sequential implementation of ID algorithms\label{alg:id-algo-application}}
\begin{algorithmic}[1]
  \Let{$msg$}{$network$} \Comment{all observed messages}
  \ForEach{$algorithm \in algorithms$}
    \State \Call{algorithm.progress\_message}{msg}
  \EndFor
  \State \dots
  \Comment{at a given interval}
  \ForEach{$algorithm \in algorithms$}
    \State \Call{algorithm.do\_housekeeping}{}
  \EndFor
\end{algorithmic}
\end{figure}

@@

\begin{figure}[H]
\captionof{lstlisting}{ID Algorithm Pattern\label{alg:id-algo-pattern}}
\begin{algorithmic}[1]
  \Require{nodes, global storage for information about nodes}
  \Function{process\_message}{$msg$}
    \ForEach{$byte \in msg$} \label{alg:id-algo-pattern-loop1}
     \State \dots \Comment{analyse byte-sequences}
    \EndFor
    \Let{$nodes_x$}{value}  \Comment{optionally update node info}
    \State \Call{send}{$nodes_y$, ``info''} \Comment{optionally exchange info} \label{alg:id-algo-pattern-send1}
  \EndFunction
  \State
  \Function{do\_housekeeping}{}
    \ForEach{$node \in nodes$} \label{alg:id-algo-pattern-loop2} \label{alg:id-algo-pattern-common-data}
      \If{$node > \dots$} \Comment{validate recorded value}
        \State \dots \Comment{take actions}
        \State \Call{send}{$node$, ``info''} \Comment{e.g. communicate} \label{alg:id-algo-pattern-send2}
      \EndIf
    \EndFor
  \EndFunction
  \State
\end{algorithmic}
\end{figure}

@@

To bridge the gap between the research world and the heterogeneous IoT where
developers want to implement an actual IDS, while still respecting the limited
resources they have at their disposal, we introduce \NAME, a framework
consisting of a DSL, code generator and software library.

The DSL enables researchers to formally describe their ID algorithms, without
focusing on any specific language or platform. Given such formal descriptions
of ID algorithms, a code generator is now able to combine these, while avoiding
the sequential patterns/problems presented in listings
\ref{alg:id-algo-application} and \ref{alg:id-algo-pattern}. It can also target
different languages and platforms, while reusing a software library that
provides e.g. message parsing, network access wrapping,\dots

\subsection{DSL}

The DSL component of \NAME has two major goals and one minor: (1) allow the
description of ID algorithms in a formal way and (2) avoid that those
descriptions contain \emph{non-fusible} logic and/or data. Besides these (3)
the usability of the language is also taken into account.

The first goal is covered by the construction of the language itself. Given a
description of an ID algorithm using the DSL, the algorithm can be
unambiguously interpreted. The DSL borrows a lot of its syntax from C-like
languages, dropping only a small subset of its capabilities, but replacing
those with other higher-level constructs taken from object-oriented (OO)
languages, aspect-orientation,\dots

The most important omission from the C syntax are basic loops. Loops are very
generic and make it hard to identify if they deal with the same logic and/or
data. Fusing these loops across different algorithms is often non-trivial. In
\NAME's DSL, other constructs have been introduced to lift the definition of
these loops to a more functional level: scheduling and scoping of the execution
of functions, reacting to events and pattern matching are three examples that
allow describing the actual meaning of loops in stead of simply the loop
itself. It sometimes requires rethinking the reason behind a technical loop,
but often results in a much clearer description.

To deal with common data, the domain provides the notion of \emph{nodes}. As
most, if not all, ID algorithms targeting the IoT require some notion of the
other nodes or devices in the vicinity, the DSL makes this the primary
information carrier. All functional constructs, mentioned above, typically
operate on the collection of known nodes, or a functionally defined subset of
that. Each algorithm also has the possibility to \emph{extend} nodes, adding
its own properties needed for its execution.

Based on a C-like syntax the language is also very accessible. To further
nurture the usability, syntactic sugar from OO languages is also adopted.
Because not everything should be implemented in such a high-level language, it
also provides a way to \emph{import} external functionality. In the first place
this will include a standard library of functions that can be called, but it
can include any custom built software component. Finally it tries to limit the
need for type information and performs as much type inference as possible. As
it aims to be language that describes algorithms on a functional level it seeks
to safeguard its users from these \emph{technicalities}. The fact that the
opportunities of such a formal description include generation of code in a
variety of languages, each with possibly different types, also results the need
for a more functional approach to types.

Listing \ref{lst:heartbeat} illustrates some of the features mentioned above,
using a small but representative example of an algorithm: a heartbeat. Using a
heartbeat, nodes can broadcast a message at a regular interval, allowing other
nodes to monitor its availability. Missing heartbeats could indicate problems,
such as physical capturing of a certain node.

\lstinputlisting[
  float,
  basicstyle=\small\ttfamily,
  label=lst:heartbeat,
  caption=Heartbeat in \NAME
]{src/heartbeat.foo}

\subsection{Code Generator}

\TODO

- semantic model \cite{fowler2010domain}
- model transformations
- code model + syntactic mapping using coding pattern-like instructions

\subsection{Software Library}

\TODO

- incoming message parsing
- outgoing message grouping
- execution of schedules
- standard library: high-level conceptual functions (e.g. now(), sha1,...)
